---
title: "Preprocess result"
author: "Yutong Liu"
date: "4/24/2019"
output: html_document
---
```{r}
library(devtools)
load_all("/Users/liuyutong/Desktop/BIOS 735/final_project/UmbrellaAcademy/UmbrellaAcademy")
```

** Pre-process **
We start by loading our package "UmbrellaAcademy",
```{r}
library(UmbrellaAcademy)
```
Next, we start by reading in the original files with function `fread` from package `data.table`.  

```{r}
setwd("~/Desktop/BIOS 735/final_project/UmbrellaAcademy/UmbrellaAcademy/data-raw")
train.data   <- fread("train.csv")
test         <- fread("test.csv")
test_labels  <- fread("test_labels.csv")
```
  
We will then take a look at the dimension and structure of the raw dataset.  
  
```{r}
dim(train.data)
train.data[1:10,]
dim(test)
test[1:10,]
dim(test_labels)
test_labels[1:10,]
```
   
Here, for the testing dataset, a lable of -1 means that this comment will not be used in the evaluation, so we delete those records from the data set.
```{r}
# exclude text_label = -1
test_labels_use <- test_labels[test_labels$toxic!=-1,]
test_use <- test[test_labels$toxic!=-1,]
```

Then, for the training set, we down sample the nontoxic comments and obtain a balanced training set.  

```{r}
# undersample nontoxic group
toxic.id    = which(train.data$toxic==1) # # of toxic = 15294
nontoxic.id = which(train.data$toxic==0) # # of nontoxic = 144277
set.seed(123)
nontoxic.sampleId = sample(nontoxic.id,length(toxic.id))
train.sample      = train.data[c(toxic.id,nontoxic.sampleId),]
```
   
We then process the training set and test set by using the function `preprocess` from our package, which will gives us the document-term matrices as the output. This function will help performing tokenization, stemming, removing stop words, removing features that appear less than 10 times across all comments and applying count vectorization.

```{r}
data.pre <- preprocess(traindata = train.sample, testdata = test_use, 
                                prep_fun = tolower, tok_fun = tok_fun)
train.dtm <- data.pre$train.dtm
test.dtm  <- data.pre$test.dtm
```
  
Let's then take a look at the the document-term matrices:
```{r}
train.dtm[1:10, 13775:13780]
test.dtm[1:10,1:4]
```
